Should I Airbnb My Property?

### 2017-01-10 abbReturns_Writeup.Rmd
Review of initial 80%.

---
title: "Should I AirBNB My Property?"
author: "Andy Krause and Gideon Aschwanden"
date: "December 17, 2016"
output:
  pdf_document: default
  html_document: default
geometry: margin=3cm
---

@@

@@

### Scenario

We approach answering the question of Airbnb-ing or not by examining a scenario likely to be faced by a recent purchaser of a property.  More specifically, we examine the following case:

* An investor purchases a property in Melbourne on Sept 1, 2015.
* They are looking to maximize the revenue over the 12 month period from Sept 1 2015 to August 31 2016.
* The choices are 1) traditional year long rental (long term), or: 2) short-term, Airbnb type rental.
* Only direct monetary costs are considered, others such as liquidity and inconvience are not considered.

We will not, a priori, specify a location or property type as we hypothesize that the feasibility of a short term rental will vary by these characteristics.  Australia has a tax policy, Negative Gearing, that does allow for favourable tax treatment to real estate investors who experience greater costs than revenues on investment properties.  However, as this policy requires knowledge of the investor's personal income level, we will not be considering this policy in our analysis here.

One benefit of a short-term rental may be the ability to adjust the nightly rate over the course of the year to reflect changing demand and market trends.  As the future trends are unknown at the time of purchase and the decision of long vs short term rental, we make our comparisons using a fixed nightly short term rate to respresent the most likely state of knowledge for our hypothetical investor on September 1, 2015. Both long and short term rates and occupancy structures will be baesd on the market conditions as of this date.

### Equation

In this above scenario we compare the 12-month returns from the a long term rental with those of a short term rental.  We define returns here as the gross revenue minus the direct costs associated with each of the two tenure possibilities.

\begin{equation} 
R_{LT} = Rev_{LT} - C_{LT}
\end{equation}

\begin{equation} 
R_{ST} = Rev_{ST} - C_{ST}
\end{equation}

\begin{equation} 
Rev_{LT} = Rent_{M} * (12 - C_{S}) 
\end{equation}

where $C_{S}$ are search costs defined in terms of the time (in months) spent to find a long-term tenant.

\begin{equation} 
Rev_{ST} = Rate_{N} * O_{Total} 
\end{equation}

where $O_{Total}$ are the total number of occupancies over the 12-month period.

\begin{equation} 
C_{LT} = T + M 
\end{equation}

where

where $T$ are tax related expenses and $M$ are general maintenance expenses.

\begin{equation} 
C_{ST} = T + M + J 
\end{equation}

where $T$ are tax related expenses, $M$ are general maintenance expenses and $J$ are janitorial expenses incurred after each short term tenancy.

We assume for the sake of this scenario that the tax $T$ and general maintenance $M$ expenses are identical between tenure options and therefore cancel out.  With the analysis that follows we will consider short term rental to be prefable to long-term if the following inequality holds:

\begin{equation} 
Rate_{N} * O_{Total} - J > Rent_{M} * (12 - C_{S}) 
\end{equation}

@@

Data for this studies were gathered from two different sources, one for short-term rental market and one for the long-term rental market.  The Airbnb online portal is used as evidence of the short-term rental market.  Airbnb is the largest short-term rental hub in the Melbourne region, though others do exist.  As opposed to similar sites such as VRBO.com and homeaway.com which focus on vacation homes in tourist dominated areas such as beaches and ski resorts, Airbnb offers considerable coverage in urban areas and, in way ways, directly compete with more traditional long term rentals.  The data on the AirBNB market in Melbourne have been purchased from www.airdna.co, a data provider specializing in AirBNB data collection and analysis. The short term data include two types of observations, property level and nightly.  Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year.  This data indicate the advertised price, the occupancy status and a unique reservation identification (useful for determining the length of stay).

The Airbnb data cover the Melbourne metropolitan region from October 1st 2014 to August 30th 2016. There are missing nightly observations during the time period before September 1 2015.  While we initially imputed these observations we found the results unreliable and have structured the analysis to only cover the time period containing complete data.  All monetary values in the Airbnb data are in US dollars and have been converted to Australian dollars using the median of the exchange rate during this time period, 1.32 Australian dollars to 1 US dollar.

The data on long-term leases was provided by Australia Property Monitors (APM), a Fairfax Group company. This data also includes an observation for each rental property each time a detail of the listing changed.  For example, if a property were listed for rent for one month and the price was dropped once each week there would be four observations in the dataset. These four observations are grouped by an event-specific identification number. In each case the last observation signifies the final price of the rental contract.  We separate this data into two sets for efficiency reasons. The first data set includes all information about the property, physical and economic (date and amount of rent contract), while the second contains a property ID, a rental event ID, the list price and the date of the change.  As such, the first dataset is a set of market observations with associated characteristics and the second is a listing history of each event. The long-term lease data covers the period from January 2014 to December 2015.

Unfortunately, there are no common identifiers between the two dataset as the Airbnb data does not include address.  Both sets of data do include latitude and longitude, but in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on a two-dimensional coordinates.  As a result of this, we are unable to directly match observations from the two datasets.  However, the number of properties that rented long term for the 2014 to 2015 period and then switched to Airbnb for the 2015 to 2016 period are likely very small and this inability to directly match properties does not greatly impact our modeling strategy.

@@

In this section we describe the operations that we have performed to prepare the two sets of raw data for analysis.  Primarily, this involves the filtering out of data errors and/or outlier values.  However, as our data are from two different sources, number of fields also require standardization so that comparisons between the two tenure options can be made.

We begin by filtering observations by time. Our hypothetical simulated decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long term rental versus an AirBNB, short term rental approach to generating income. To determine the likely long term rental rate for any given property, we will use the observed rental transactions from the market from September 1, 2014 to August 31, 2015. Taking the AirBNB approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors).  In this simulated example we will use AirBnB data from September 1, 2015 to August 31, 2016 to represent the actual AirBNB market conditions over the time period in question.

Next, we examine property types. AirBNB units are classified into 19 different property types. The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt^[Property types are owner/lister defined, therfore it is likely that the Igloo, at the very least, is not a completely true representation of the property type]. For the sake of this analysis, we will collapse these twenty types into three types:

@@

Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the AirBNB data conforms to North American lexicon.  This is most apparent when talking about apartment dwellings.  In Australia these are referred to as Units, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure).  Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, likely, Townhomes in the North American context.

As the chart below show, the **Other** category only makes up a very small percentage of the properties in both the short and long term market, especially in the long-term market.  Additionally, a good portion of the **Others** in the Airbnb data are Bed & Breakfast units, a use which doesn't fit with our research question.  As a result, we remove the 'Others' from both the Airbnb and the long term rental data.

@@

```{r type_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3, fig.cap='Tenure Type by Property Type'}

@@

Next, for the short term rentals properties can be listed as one of three types relating to the extent of the property which is able to be booked:

@@

As long-term rentals do not offer **Private Room** or **Shared Room** options^[Our dataset does not include rooming houses, purpose built student accomodations and other types of rental which may offer private or shared room options] and our purpose here is a comparison of long and short term returns, we remove all short term properties that do not lease the entire home/apt. Unfortunately, this filter does remove about 40% of the short term data, however, given the research question it is unavoidable.

```{r ltype_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.width=5, fig.cap='Listing Type of Short Term Rentals'}

@@

  ggplot(abb.data, aes(x=listing.type)) + 
    geom_bar(color=abb.col[1], fill=abb.col[1]) +
    xlab('') + 
    ylab('# of Properties\n') +
    labs(title = "Short Term Rentals by Listing Type\n") + 
    theme(plot.title = element_text(hjust = 0.5))

@@

We begin by removing all properties that are missing bedroom information or have a likely data error in this fields (more than 14 bedrooms). For both short and long term rentals we see that most apartments have less than 4 bedrooms and houses less than 5.  We filter the data accordingly. Additionally, note that short term rentals consider studio apartments to be 0 bed units while the long term data considers these as 1 bedroom. We convert the 0 bed short term units to 1 bed units.

@@

We then turn to bathrooms, starting by looking at the distribution of baths between short/long and apt/house. Here we see that nearly all properties have 3 or fewer bathrooms and we filter the data accordingly. A few short term properties state no baths, which is likely an error so we remove these observations from the data. Also, here we notice that the short term data gives bathrooms in halves, while the long term data presents baths in whole numbers.  We round the baths in the short term data up to whole integers.

@@

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short term rental) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two.

@@

  stb <- ggplot(abb.data, aes(x=bedbath)) +
    geom_bar(fill=str.col, color=str.col) + 
    xlab('') +
    ylab('# of Props\n') +
    labs(title = "Bed/Bath Combo in Short Term Rentals\n") + 
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme(axis.text.x = element_text(angle = 90, vjust=0.4)) +
    facet_wrap(~type)

@@

From this analysis we see that there are 6 combinations that make up most all properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2 and 4/2. We filter the data to these combinations.

@@

We now turn to filtering by the prices -- the nightly rates and weekly rental values. The long term rental data have a single price observation, the asking rent each time an change in the listing occurs, including the price at closing.  For the short term data, there are two indicators of nightly rate.  First, each property has a average nightly rate figure that takes the average of the rates for all booked nights over the October 2014 to August 2016 period. Second, there is a rate for each day in the daily observations. For the vast majority of properties, the daily rates change very little, but for a select few there is a wide variation in the daily rates, often somewhat randomly.  In cases where a small number of reservations have been made, the average nightly rate can deviate widely from usual rate due to outliers on the high end. To remedy this, we create a new property-level variable that is the median nightly rate of successful bookings over the September 1 2015 to August 31, 2016 period.

We start by removing property level observations missing price data. Plotting the distribution of the rates and rents we see that most long term weekly rentals are greater than \$200 and less than \$1000 per week while most median nightly rates are greater than \$50 and less than \$500 per night. We filter the property level data accordingly.  We also filter the short term daily data to include only those observations with nightly rates between \$50 and \$500.

@@

Location is clearly an important determinant in both the short and long term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs.  In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts.  The specific suburb of a property is the second, or finer scale, at which the market operates. The plot below shows the three large submarkets (Inner, Middle and Outer) and the individual suburbs boundaries.

@@

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analysing short and long term rentals in Melbourne:

@@

We select the 18 suburbs below to represent the 'core' of the city.  The majority of tourist destinations and major events such as the Australian Open and the Grand Prix are located in these suburbs.  All are well serviced by public transportation and possess abundant amenties for tourists.

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short and long term data. Finally, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs). The location of the short and long term rentals, colored by submarket, are shown below

Before assigning submarkets to the long term property we must remedy the fact that about 5% of the long term rental observations have missing latitude and longitude values.  Half of these do, however, have a lat/long value for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps for those properties that lack a specific lat/long value we apply the street centroid value.  For those without either we remove them from the dataset at this point. After making this correction we then assign one of the five submarket designations based on the process described above.

@@

```{r loc.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=5.5, fig.height=9}

@@

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here.  First, the vast majority of Airbnb units are apartments in city-core or city locations.  As expected and as shown by the map, the long term properties are more evenly spread across the metro region than the short term properties.  Long term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short term market, especially in apartments, while suburban and rural apartment, somewhat common in long term markets, are nearly absent in the Airbnb market.  Overall, the suburban and rural Airbnb market are quite thin, later results focusing on these market will need to be examined in light of this small sample size.

@@

```{r abbtype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=9,fig.height=3.5}

  abbtype.map

@@

```{r ltrtype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=9,fig.height=3.5}

  ltrtype.map

@@

### Actual Revenue

For the short term data, we first calculate the total number of bookings, the occupancy rate, blocked rate and proportion of the year over which the property was listed.  Using this data we sum up the total revenue from bookings. For properties that were not listed on the site for the entire year, we impute their annual revenue by dividing the actual revenue by the proportion of the year in which they were listed.^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] Properties with no reserved bookings, and, therefore, no revenue, are removed.

Looking at the distribution of the revenues from each tenure type, we see that those for the long-term rentals are for more consistent, with the vast majority randing between \$12,000 to \$30,000 per year.  Short term revenue range from neglible to highs of greater than \$80,000.  While this simple analysis does not control for locational and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not.  Also, the high number of very low yield Airbnb properties also suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.

```{r load_workspace2, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE}

 load(paste0(data.path, '/analyzed/abb_results.RData'))

```{r revdensity.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5}

@@

We have applied nine separate filters to our two datasets.  In the name of transparency and to highlight the bounds of our research questions it is useful to review this filters. There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Removing those not listed during the study period removes 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 6,600 properties structural characteristics, locational issues and outlying nightly rates remove another 800 from consideration.  Finally, 1,400 of these have never been booked and have \$0 revenue, requiring removal.  From over 20,000, we have 4,446 Airbnb properties that meet the requirements of our research question.

For the long term data, over half of the observations fall outside of our considered time frame,reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 10,000 observations, leaving 122,051 in the final dataset.

>>>>

---
title: "Should I Airbnb My Property?"
author: "Andy Krause and Gideon Aschwanden"
date: "January 9, 2017"
output:
  pdf_document: default
  html_document: default
geometry: margin=3.4cm
bibliography: abb.bib
---

@@

 ## Set options

  options(xtable.comment = FALSE)

@@

We approach answering the question of whether to pursue a short- or long-term leasing strategy by creating and assessing a scenario likely to be faced by a recent purchaser of a property.  More specifically, we examine the following case:

* An investor purchases a property in Melbourne on Sept 1, 2015.
* They are looking to maximize the revenue over the 12-month period from September 1, 2015 to August 31, 2016.
* The choices are 1) traditional year long rental (long term), or: 2) short-term, Airbnb type rental.
* Only direct monetary revenues and costs are considered, others such as liquidity and inconvience are not considered.

We will not, a priori, specify a location or property type as we hypothesize that the preference of a short-term rental over a long-term one will vary by the location and property characteristics.  Additionally, Australia implements a particular tax policy, often termed `negative gearing', that does allow for favourable tax treatment to real estate investors who experience greater costs than revenues on investment properties^[See @fane2004negative for a complete explanation of negative gearing.  However, as the negative gearing policy requires knowledge of the investor's personal income situation, we will not be considering any effects of this policy in our analysis here.

One benefit of a short-term rental may be the ability to adjust the nightly rate over the course of the year to reflect changing demand and market trends. As future market trends are unknown at the time of purchase and the decision of long- vs short-term leasing strategy is asymetrically irreversable (can go from short to long, but not long to short), we make our comparisons using a fixed nightly (short-term) rate to respresent the most likely state of knowledge for our hypothetical investor on September 1, 2015. Both long- term rents and short-term rates and occupancy structures will be baesd on the market conditions as of this date.

We do not directly observe properties that are utilized in both a short and long term leasing strategy.  Therefore, comparisons are done based on estimated or imputed revenues.  Most specifically, we will estimate short-term revenue for observations of long-term rentals (LTRs) and long-term revenue for observations of short-term rentals (STRs). A preference for short or long term leasing strategy will be determined based on comparing the observed and estimated revenues.

### Comparison

In this above scenario we compare the 12-month revenues from the a long-term rental with those of a short-term rental.  We define returns here as the gross revenue minus the direct costs associated with each of the two tenure possibilities.

\begin{equation}
Rev_{LT} = RevG_{LT} - C_{LT}
\end{equation}

\begin{equation}
Rev_{ST} = RevG_{ST} - C_{ST}
\end{equation}

Where $Rev_{LT}$ is the net revenue for a long-term rental, $RevG_{LT}$ is the gross revenue and $C_{LT}$ are the costs associated with long-term rental. The gross long-term revenue is expressed by:

\begin{equation}
RevG_{LT} = Rent_{M} * (12 - C_{S})
\end{equation}

where $Rent_{M}$ is the nominal monthly rent and $C_{S}$ are search costs defined in terms of the time (in months) spent to find a long-term tenant.

The short-term revenue is expresse by:

\begin{equation}
RevG_{ST} = Rate_{N} * O_{Total}
\end{equation}

where $Rate_{N}$ is the average nightly rate and $O_{Total}$ are the total number of occupancies (bookings) over the 12-month period.

We have limited our consideration of costs in this case to be those undergone during a single year of ownership.  Broadly speaking, these fall into three categories for long-term rental and four categories for short-term properties.  Long-term costs are summarized as:

\begin{equation}
C_{LT} = T + M + F
\end{equation}

where $T$ are tax related expenses, $M$ are general maintenance expenses and $F$ are finaning costs. Short-term properies experience the same three costs ($T$, $M$ and $F$) plus additional janitorial costs between each of the short-term occupants.

\begin{equation}
C_{ST} = T + M + F + J
\end{equation}

As our scenario compares a hypothetical and particular owner over the course of a single  year the tax $T$ and $F$ costs are identical between the two leasing strategy and therefore cancel out. General maintenance $M$ expenses may differ between the two strategies, and an argument could be made for either to be higher. Long-term renters spend more time in the property over the course of the year and bring their own appliance and furniture which may damage the home. They are also more likely to affix items to the structure.  Conversely, short-term rentals are usually furnished and the occupants have less incentive not to be evicted as their damage is not often uncovered until after they have left^[Note that Airbnb's rating policy may help protect hosts as bad guests can be flagged on the platform.].  Both arguments are meritous and we were unable to find any evidence to support maintenance costs being higher in short or long term rentals. As a result we have made the assumption that the maintenance costs will be identical, and they too cancel out.

Finally, we have janitorial costs $J$. These cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experiences every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform.  In addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb. Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior.  As a result of janitorial fees being a separate line item and directly paid by the occupants can also be ignored in the comparative analysis below as they are direct.

Given that the costs cancel out between the two options, we will consider short-term rental (STR) strategy to be prefable to long-term rental (LTR) strategy if the following inequality holds:

\begin{equation}
RevG_{ST} > RevG_{LT} 
\end{equation}

More specifically, if:

\begin{equation}
Rate_{N} * O_{Total} > Rent_{M} * (12 - C_{S})
\end{equation}

@@

Data for this studies were gathered from two different sources, one for short-term rental market and one for the long-term rental market.  Short-term rental market observations are from the Airbnb online portal. Airbnb is the largest short-term rental hub in the Melbourne region. As opposed to similar sites such as VRBO.com and homeaway.com which focus on vacation homes in tourist dominated areas such as beaches and ski resorts, Airbnb offers considerable coverage in urban areas and, therefore, directly competes to a greater degree with more traditional long-term rentals than the other sites. The data on the AirBNB market in Melbourne have been purchased from www.airdna.co, a data provider specializing in AirBNB data collection and analysis. The short term data include two types of observations, property level and nightly. Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year during its time on the portal.  These data indicate the advertised price, the occupancy status and a unique reservation identification (useful for determining the length of stay).

The Airbnb data cover the Melbourne metropolitan region from October 1st, 2014 to August 30th, 2016. There are missing nightly observations during the time period before September 1, 2015. After initially imputed these observations we found the results unreliable and have structured the analysis to only cover the time period containing complete data.  All monetary values in the Airbnb data are in US dollars and have been converted to Australian dollars using the median of the exchange rate during this time period, 1.32 Australian dollars to 1 US dollar.

The data on long-term leases was provided by Australia Property Monitors (APM), a Fairfax Group company. In raw form, this dataset includes an observation for each rental property listing for each time a detail of the listing changed. For example, if a property were listed for rent for one month and the price was dropped once each week there would be four observations in the dataset. These four observations are grouped by an event-specific identification number. This number is specific to the listing event, not the property. In each case the last observation signifies the final price of the rental contract. We separate this data into two sets for efficiency reasons. The first data set includes all information about the property, physical and economic (date and amount of rental contract), while the second contains a property ID, a rental event ID, the list price and the date of the change. As such, the first dataset is a set of market observations with associated characteristics and the second is a listing history of each rental event. The long-term lease data covers the period from January 2014 to December 2015.

Unfortunately, there are no common identifiers between the two datasets as the Airbnb data does not include address. Both sets of data do include latitude and longitude, but in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on a two-dimensional coordinates. As a result of this, we are unable to directly match observations from the two datasets. However, the number of properties that rented long term for the 2014 to 2015 period and then switched to Airbnb for the 2015 to 2016 period are likely very small and this inability to directly match properties does not greatly impact our modeling strategy.

@@

In this section we describe the operations that we have performed to prepare the two sets of raw data for analysis.  Primarily, this involves the filtering out data errors, outlier values and other observations than contain data that do not fit our research questions. Additionally, as our data are from two different sources, a number of fields also require standardization so that comparisons between the two leasing strategy options  -- long-term rental (LTR) vs short-term rental (STR) -- can be made.

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an AirBNB-type, short-term rental approach to generating income. To determine the likely LTR rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use AirBnB data from September 1, 2015 to August 31, 2016 to represent the actual AirBNB market conditions over the time period in question.

Next, we filter the data based on property types. AirBNB units are classified into 19 different property types. The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt^[Property types are owner/lister defined, therfore it is likely that some designation, such as Igloo, at the very least, are not completely true representations of the property type]. For the sake of this analysis, we will collapse these twenty types into three types:

@@

Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the short-term data conforms to North American lexicon.  This is most apparent when talking about apartment dwellings.  In Australia these are referred to as Units, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure)^[Also note that 'Apartment-type' ownership is very rare in Australia.  Nearly every multi-family dwelling is owned as a condominium.  See @Berry2000investment for more information on the overall structure of the Australian market]. Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, likely, Townhomes in the North American context.

The **Other** category only makes up a very small percentage of the properties in both the short and long term market; especially so in the long-term market (Figure 1) .  Additionally, a large proportion of the **Others** in the short-term data are Bed & Breakfast units, a use which doesn't fit with our research question. As a result, we remove the 'Others' from both the short- and the long-term rental data.

@@

```{r type_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3, fig.cap='Tenure Type by Property Type', fig.lp='typecomp'}

@@

Next, the short-term rentals properties can be listed as one of three types relating to the extent of the property which is able to be booked:

@@

As long-term rentals do not offer **Private Room** or **Shared Room** options^[Our dataset does not include rooming houses, purpose built student accommodations and other types of rentals which may offer private or shared room options] and our purpose here is a comparison of long- and short-term returns, we remove all short-term properties that do not lease the entire home or apartment. Unfortunately, this filter does remove about 40% of the short-term data (See Figure 2), however, given the research question it is unavoidable.

```{r ltype_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=2.5, fig.width=3.75, fig.cap='Listing Type of Short Term Rentals'}

@@

  ggplot(abb.data, aes(x=listing.type)) +
    geom_bar(color=abb.col[1], fill=abb.col[1]) +
    xlab('') +
    ylab('# of Properties\n') +
    labs(title = "Short Term Rentals \nby Listing Type\n") +
    theme(plot.title = element_text(hjust = 0.5))

@@

We begin by removing all properties that are missing bedroom information or have a likely data error in this fields (more than 14 bedrooms). For both short and long term rentals we see that most apartments have less than 4 bedrooms and houses less than 5 (Figure 3).  We filter the data accordingly. Additionally, note that short term rentals consider studio apartments to be 0 bed units while the long term data considers these as 1 bedroom. We convert the 0 bed short term units to 1 bed units.

@@

We then turn to bathrooms, starting by looking at the distribution of baths between short/long and apartment/house. Here we see that nearly all properties have 3 or fewer bathrooms and we filter the data accordingly (Figure 4). A few short term properties indicate no baths, which is likely an error so we remove these observations from the data. Also, here we notice that the short term data gives bathrooms in halves, while the long term data presents baths in whole numbers.  We round the baths in the short term data up to whole integers.

@@

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short-term rental) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two (Figure 5). From this analysis we see that there are 6 combinations that make up most all properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2 and 4/2. We filter the data to these combinations.

@@

  stb <- ggplot(abb.data, aes(x=bedbath)) +
    geom_bar(fill=str.col, color=str.col) +
    xlab('') +
    ylab('# of Props\n') +
    labs(title = "Bed/Bath Combo in Short Term Rentals\n") +
    scale_y_continuous(breaks=c(1000, 3000, 5000)) +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(axis.text.x = element_text(angle = 90, vjust=0.4)) +
    facet_wrap(~type)

@@

@@

Next we filter by the prices -- the nightly rates and weekly rental values. The long term rental data have a single price observation, the asking rent at closing.  For the short term data, there are two indicators of nightly rate.  First, each property has a average nightly rate figure that takes the average of the rates for all booked nights over the October 2014 to August 2016 period. Second, there is a rate for each day in the daily observations. For the vast majority of properties, the daily rates change very little, but for a select few there is a wide variation in the daily rates, often somewhat randomly. In cases where a small number of reservations have been made, the average nightly rate can deviate widely from usual rate due to outliers on the high end. To remedy this, we create a new property-level variable that is the median nightly rate of successful bookings over the September 1, 2015 to August 31, 2016 period.

We start by removing property level observations missing price data. Plotting the distribution of the rates and rents we see that most long term weekly rentals are greater than \$200 and less than \$1000 per week while most median nightly rates are greater than \$50 and less than \$500 per night (Figure 6). We filter the property level data accordingly.  We also filter the short term daily data to include only those observations with nightly rates between \$50 and \$500.

@@

Location is an important determinant in both the short- and long-term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs.  In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts. The specific suburb of a property is the second, or finer scale, at which the market operates. Figure 7 shows the three large submarkets (Inner, Middle and Outer) and the individual suburbs boundaries.

@@

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analyzing short and long term rentals in Melbourne:

@@

We select the 18 suburbs below to represent the 'core' of the city (Table 1).  The majority of tourist destinations and major events such as the Australian Open and the Grand Prix are located in these suburbs.  All are well serviced by public transportation and possess abundant amenities for tourists.

```{r, results="asis", comment=FALSE, message=FALSE, echo=FALSE}

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short and long term data. Finally, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs). The location of the short and long term rentals, colored by submarket, are shown in Figure 8.

Before assigning submarkets to the long term property we must remedy the fact that about 5% of the long term rental observations have missing latitude and longitude values.  Half of these do, however, have a latitude and longitudeg value for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps we apply the street centroid value to the observations with missing lat/long values.  For those without either set of lat/longs we remove them from the dataset at this point. After making this correction we then assign one of the five submarket designations based on the process described above.

@@

```{r loc.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=5.5, fig.height=9, fig.cap='Location Maps'}

@@

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (Figure 9). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long term properties are more evenly spread across the metro region than the short term properties.  Long term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short term market, especially in apartments, while suburban and rural apartments, somewhat common in long term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

```{r abbtype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=9,fig.height=7, fig.cap='Observations by Submarket'}

  ggMultiPlots(abbtype.map, ltrtype.map, cols=1)

@@

### Observed Revenue

For the short-term data, we first calculate the total number of bookings, the occupancy rate, blocked rate and proportion of the year over which the property was listed.  Using this data we sum up the total revenue from bookings. For properties that were not listed on the site for the entire year, we impute their annual revenue by dividing the actual revenue by the proportion of the year in which they were listed.^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] Properties with no reserved bookings, and, therefore, no revenue, are removed.

Looking at the distribution of the revenues from each tenure type, we see that those for the long-term rentals are for more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year (Figure 10).  Short term revenues range from neglible to highs of greater than \$80,000.  While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not. Also, the high number of very low yield Airbnb properties also suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.

```{r load_workspace2, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, cache=FALSE}

 load(paste0(data.path, '/analyzed/abb_results.RData'))
 a<-1

@@

```{r revdensity.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='Comparison of Revenue Distributions'}

@@

We have applied nine separate filters to our two datasets.  In the name of transparency and to highlight the bounds of our research questions it is useful to review this filters (Table 2). There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Filtering those not listed during the study period removes 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 6,600 properties structural characteristics, locational issues and outlying nightly rates remove another 800 from consideration.  Finally, 1,400 of these have never been booked and have \$0 revenue, requiring removal.  From over 20,000, we have 4,446 Airbnb properties that meet the requirements of our research question.

For the long term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 10,000 observations, leaving 122,051 in the final dataset.
